Solar wind, F10.7, and geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit

### 2016-10-17 paper/paper.tex
First pass of paper revisions

\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \cite{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.  When hourly averages are considered, a significant peak is found to occur six hours after event onset, and the primary factor that determines the post-onset peak amplitude in $\rho_{eq}$ is elevated $F10.7$.  In addition, for hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event depends on the north-south component of the interplanetary magnetic field, $B_z$ after the time of onset, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ value 7-11 hours after onset.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

The dependence of the plasmapause location on geomagnetic activity was observed as early as \cite{Carpenter1966} and there are well--validated models of the plasmapause location as a function of $L$, $MLT$, and geomagnetic activity (\cite{LemaireEarthsPlasmasphere}, \cite{Moldwin2002ModelPlasmapause}, \cite{OBrien2003EmpiricalPlasmapause}).  Models also exist for the plasmasphere density (\cite{Gallagher1988EmpiricalModelPlasmasphere}; \cite{LemaireEarthsPlasmasphere}) with similar parameterization.

The dependence of the plasma density outside of the nominal plasmapause location on geomagnetic and solar wind conditions has not been extensively studied until the past decade (\cite{Takahashi2006}; \cite{Takahashi2010}; \cite{Denton2016}). Earlier works that considered profiles of ion density found a significant dependence of the plasmapause location on geomagnetic activity (\cite{Chappell1970}; \cite{Maynard1977}; \cite{Carpenter1992}), but the relationship between geomagnetic activity and density outside of the plasmapause was not explicitly considered.

\cite{Takahashi2006} estimated magnetospheric mass density using measurements from the CRRES satellite during a 73-day period in 1991 when CRRES was in an elliptical, near equatorial orbit and primarily beyond the nominal plasmapause. They found that the local average ion mass, $M$, had some correlation with geomagnetic activity, with more negative hourly $D_{st}$ values corresponding to higher $M$ and larger 1.5- and 3-day averages of $K_p$ corresponding to higher $M$ (3-hour averages of $K_p$ had little visual correlation with $M$). The mass density estimates were obtained from CRRES observations in the MLT range of [12:00, 18:00] with most observations at $L$ between 5 and 7 $R_E$.

\cite{Denton2006} found a weak relationship between $D_{st}$ and $K_p$ and the field line distribution of plasma mass density, $\rho$, using observations of the first three toroidal Alfv\'en harmonic frequencies for $L$ in the range of 6-8~$R_E$.  A more pronounced peak in the distribution along a field line near the equator was found for lower $D_{st}$ and larger $K_p$, but the equatorial value of $\rho$ was found to be similar for $D_{st}$ bins of [-142, -31] and [-31, 37] nT and $K_p$ bins of [1.5,3.4] and [3.4, 5.9].

\cite{Takahashi2010} developed a mass density dataset using measurements from the Space Environment Monitor instruments on the Geostationary Operational Environmental Satellites (GOES) satellites from 1980 through 1992, with most measurements in the range of $L=6.8\pm0.2 R_E$. The mass density was estimated using the Alfv\'en wave velocity relationship, $V_A=B/\sqrt{\mu_0\rho}$, a magnetic field model, and a numerical solution to a wave equation with an ionospheric boundary and the assumption of a zero resistance ionosphere.  The equatorial mass density, $\rho_{eq}$, was derived from the estimated mass density using a power law dependence on the geocentric distance to the field line of the observation, $R$, $\rho=\rho_{eq}(LR_E/R)^{1/2}$.  They found a high correlation ($\sim 0.94$) between 27-day averages of $F10.7$ and 27-day medians of $\rho_{eq}$.

\cite{Takahashi2010} noted that downward drops in the $D_{st}$ index coincided with significant changes in $\rho_{eq}$ for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ increases after $D_{st}$ minimum, two had $\rho_{eq}$ spikes before the minimum, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, the epoch average of $D_{st}$ for storms with a minimum $D_{st} < -50$~nT showed an enhancement in $\rho_{eq}$ the same day as minimum $D_{st}$ in the epoch averages.

\cite{Yao2008} studied the relationship between $D_{st}$ and the number density of low-energy $O^+$ in different regions (ring current and plasma sheet) using the TEAMS and ESA instruments on the low-altitude and high-inclination orbiting FAST satellite and found that the average $N_{O^+}$ across the sampled L-shells (2-14) had a strong correlation (0.88) with the minimum $D_{st}$ of the storm.  Although the correlation at geosynchronous distances was not calculated, the data presented for four events show that the enhancements in $N_{O^+}$ tended to appear near $D_{st}$ minimum ($\pm$ 2~hours).

The above results indicate that (1) the mass density at geosynchronous distances is best correlated with $F10.7$ and (2) there exists a much weaker statistical relationship between mass density and the geomagnetic activity indices $K_p$ and $D_{st}$.  In this work we consider the dependence of mass density estimates in the \cite{Takahashi2010} data set and attempt to identify and statistically characterize the relationship between geomagnetic activity with mass density at geosynchronous distances.  In addition, we attempt to identify solar wind variables that may drive or be associated with the changes.

%There are several issues that are addressed in detail: (1) the plasma density measurements are sparse, (2) The plasma trough density has a high correlation with F10.7, and geomagnetic and solar wind parameters also correlate with F10.7, (3) the dependence of mass density on the north/south component of the solar wind magnetic field is expected to be nonlinear. For northward IMF, corresponding to low solar wind driving of the magnetosphere, the plasmasphere can expand past geosynchronous orbit where trough estimates of mass density are made resulting in an increase in observed density that are not due to solar wind driving; for southward IMF, events studies have shown an increase in mass density associated with strong southward IMF and enhanced magnetospheric activity (\cite{Takahashi2010}; \cite{Yao2008}).

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ are from the dataset of \cite{Takahashi2010} data are available from 1980 through 1991 from GOES 2, 3, 5, 6 and 7.  All other parameters used are from \cite{Kondrashov2014ReconstructionOfGaps}, which has coverage from 1972 through 2013 and are on a 1--hour time grid. $\rho_{eq}$ is the inferred equatorial mass density based on the 3rd harmonic torodial frequency of magnetic field measurements as described in \cite{Takahashi2010}.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly median over the same interval for which the solar wind parameters were averaged, the median of all $\rho_{eq}$ values in a given hour window was used. GOES 6 had the longest span of available (May 1983 to August 1991) data and the primary results presented are for GOES 6 while measurements from other satellites were used for consistency checks.

@@

The top panel of Figure~\ref{fig:ccplot} shows the long-term trends of log$_{10}(\rho_{eq})$ from GOES 6 and $F_{10.7}$ computed using the median values in 27-day non-overlapping windows.  A scatter plot of these two lines is shown in the lower panel of Figure~\ref{fig:ccplot} along with that for 1--day and 1--hour medians.  The linear correlation for GOES 6 using all measurements is found to be $0.94$, which is the same value documented by \cite{Takahashi2010} who used combined measurements from all satellites in the time interval of 1980 through 1991 using the constraint $06\mbox{:}00 \leq MLT \le 12\mbox{:}00$, $K_{p3d}\ge1.0$ and $D_{st}\ge -50$~nT.  Our 27-day correlations for GOES 2, 5, and 7 are 0.81, 0.78, and 0.87, and the time range of coverage of measurements from these satellites is 1980-01-01/1983-05-16, 1983-01-01/1987-02-28, and 1983-05-27/1991-08-29, respectively.

@@

Two types of events are considered. The first is a drop in the $D_{st}$ index below the threshold value of -50~nT. These types of events are considered in order to determine if there is a well-defined $\rho_{eq}$ dependence on geomagnetic activity as indicated by the $D_{st}$ index.  It is expected that the plasmasphere will exhibit two responses near the time of these threshold crossings.  First is the movement of the plasmapause earthward (\cite{LemaireEarthsPlasmasphere}) and the second is a possible change in density due to magnetospheric processes.

@@

Our first analysis uses GOES-6 data only from the time interval 1989-1991, which corresponds to the interval used in the epoch analysis of \cite{Takahashi2010}. In this interval, there were 75 $D_{st}$ events where $D_{st}$ remained below the $-50$~nT threshold for at least 12 hours. Only events with an onset time when the spacecraft was located between 06:00 and 12:00 MLT were considered. The first hour of a local minimum in $D_{st}$ following a threshold crossing defines the zero epoch time, and for each event, $4\cdot24$ hours were considered before and $8\cdot24$ hours after onset. For multiple detected threshold crossings within 24 hours of each other, the first crossing was selected as the event onset because visual inspection showed the subsequent crossings to typically be from brief downward deviations during the recovery period of a geomagnetic storm. To compute the epoch averages on a daily time scale, the median value of all available measurements for all events centered on a window of $\pm 12$ hours of the epoch zero hour was computed, and these averaging windows were shifted in increments of $24$~hours. Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians for each event in 1-day bins and then compute the medians across the averaged event time series on each epoch day.

The stack plot in the upper panel of Figure \ref{fig:DailyAveragedDstEvents} shows the epoch averages for the 75 events from 1989-1991 using data from GOES 6.  The minimum $D_{st}$ median is $-75$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$ on the day of the event relative to the day before and after, although the magnitude of increase observed here is $4.5$ amu/cm$^3$ instead of the increase of ${\sim} 10$ amu/cm$^3$ found in Figure 11 of \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute the medians and red error lines.  The uncertainty is shown with error lines for each parameter that have a width that is twice the standard deviation of the values used in computing the median divided by the square root of the number of values. For the non-$\rho_{eq}$\ parameters, the number of measurements used in computing the medians is typically equal to the number of events, which is much larger than the number of $\rho_{eq}$\ values as all available measurements were used instead of restricting to only values where a $\rho_{eq}$ value also existed.

@@

To increase the number of events, we have removed the constraint that the events must occur in the MLT range of 06:00-12:00 and the constraint that $D_{st}$ remained below the threshold value for 12 hours or longer.  The result is shown in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents}. The two variations from a baseline (one downwards before onset and one upwards after onset) were tested for significant deviation from two baselines, one 24-12 hours before onset, one 16-48 hours after onset. The probability of the downward variation coming from the same distribution as the data from the left baseline is 1.06\%, whereas the probability of the right upward variation coming from the same distribution of the data for the right baseline is 0.17\%.  This significant result is also maintained for GOES 2, 5, and 7.  This indicates that there are statistically significant variations in $\rho_{eq}$ within 6 hours of onset when hourly averages are considered.  The signficance of the upward variation result is consistent with previous observations of large mass density increases in the plasma trough region during large geomagnetic storms (\cite{Yao2008}; \cite{Takahashi2010}).

@@

Figure~\ref{fig:HourlyAveragedRhoEvents} can be compared with Figure~9 of \cite{Denton2016}, which shows daily averages of $\rho_{eq}$ from this dataset after the onset of quiet $K_p$ intervals.  The growth shown here (using the $\rho_m$ threshold criteria with no constraint on $K_p$) is approximately linear and lasts $\sim$ 12--18 hours whereas the 10 $K_p$ events (elevated $K_p$ followed by 2 days with $K_p \le 1^+$) considered by \cite{Denton2016} showed growth that lasted for at least 48~hours.

@@

Figure~\ref{fig:F107Binned} shows how the epoch average curve of $\rho_{eq}$ in Figure~\ref{fig:HourlyAveragedRhoEvents} depends on $F10.7$.  Statistically significant differences occur independent of whether the separating factor is the average of $F10.7$ before or after onset, as expected because the cadence of $F10.7$ is one day and its time scale of variation is longer than that of $\rho_{eq}$.  Events with high $F10.7$ are associated with higher $\rho_{eq}$ before and after event onset but similar peak values.  As was the case for $B_z$, there is an approximately 12--hour time span of linear growth of $\rho_{eq}$ near the time of onset but because the initial starting $\rho_{eq}$ is higher for high $F10.7$, the associated growth rate is lower, consistent with \cite{Denton2016}.

@@

In recent years, various works have considered the influence of solar wind control and the geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit, generally using event studies and linear correlations (\cite{Takahashi2006}; \cite{Denton2006}; \cite{Yao2008}; \cite{Takahashi2010}; \cite{Denton2011}; \cite{Denton2016}).  The event studies show a possible relationship between $\rho_{eq}$ and solar wind and geomagnetic activity indicators such as $K_p$ and $D_{st}$ while correlation studies have generally found a very weak relationship.

@@

%Based on a weak linear relationship between $\rho_{eq}$ and $B_z$, \cite{Denton2016} speculated that this may be a result of a closed magnetosphere being more conducive to plasmasphere refilling.

@@

For $\rho_{eq}$ events, more positive $B_{z}$ after onset corresponded to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may be due to the apparent mass refilling mechanism described by \cite{Denton2016}.  In order to separate this mechanism from an enhancement due to magnetospheric activity using this type of analysis, additional information about the location of the plasmapause may be required.

@@

\newpage
%\footnotesize
%\bibliographystyle{plainnat}
\bibliography{paper}

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.45]{figures/alldata-GOES6-1983-1991.eps}
\caption{Overview of data used in this article. The top four panels show parameters from \cite{Kondrashov2014ReconstructionOfGaps} and the bottom panel contains $\rho_{eq}$ based on GOES 6 measurements from \cite{Denton} after interpolation and averaging described in the text. Dashed horizontal lines in the $D_{st}$ and $\rho_{eq}$ panels indicate sample event cutoff thresholds of $D_{st} = -50$~nT and $\rho_{eq} = 20$~amu/cm$^3$ considered in Section~3.}
\label{fig:AllData}
\end{figure}

>>>>

\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \citet{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.  When hourly averages are considered, a significant peak is found to occur six hours after event onset, and the primary factor that determines the post-onset peak amplitude in $\rho_{eq}$ is elevated $F10.7$.  In addition, for hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event depends on the north-south component of the interplanetary magnetic field, $B_z$ after the time of onset, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ value 7-11 hours after onset.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

The dependence of the plasmapause location on geomagnetic activity was observed as early as \citet{Carpenter1966} and there are well--validated models of the plasmapause location as a function of $L$, $MLT$, and geomagnetic activity (\citet{LemaireEarthsPlasmasphere}, \citet{Moldwin2002ModelPlasmapause}, \citet{OBrien2003EmpiricalPlasmapause}).  Models also exist for the plasmasphere density (\citet{Gallagher1988EmpiricalModelPlasmasphere}; \citet{LemaireEarthsPlasmasphere}) with similar parameterization.

The dependence of the plasma density outside of the nominal plasmapause location on geomagnetic and solar wind conditions has not been extensively studied until the past decade (\citet{Takahashi2006}; \citet{Takahashi2010}; \citet{Denton2016}). Earlier works that considered profiles of ion density found a significant dependence of the plasmapause location on geomagnetic activity (\citet{Chappell1970}; \citet{Maynard1977}; \citet{Carpenter1992}), but the relationship between geomagnetic activity and density outside of the plasmapause was not explicitly considered.

\citet{Takahashi2006} estimated magnetospheric mass density using measurements from the CRRES satellite during a 73-day period in 1991 when CRRES was in an elliptical, near equatorial orbit and primarily beyond the nominal plasmapause. They found that the local average ion mass, $M$, had some correlation with geomagnetic activity, with more negative hourly $D_{st}$ values corresponding to higher $M$ and larger 1.5- and 3-day averages of $K_p$ corresponding to higher $M$ (3-hour averages of $K_p$ had little visual correlation with $M$). The mass density estimates were obtained from CRRES observations in the MLT range of [12:00, 18:00] with most observations at $L$ between 5 and 7 $R_E$.

\citet{Denton2006} found a weak relationship between $D_{st}$ and $K_p$ and the field line distribution of plasma mass density, $\rho$, using observations of the first three toroidal Alfv\'en harmonic frequencies for $L$ in the range of 6-8~$R_E$.  A more pronounced peak in the distribution along a field line near the equator was found for lower $D_{st}$ and larger $K_p$, but the equatorial value of $\rho$ was found to be similar for $D_{st}$ bins of [-142, -31] and [-31, 37] nT and $K_p$ bins of [1.5,3.4] and [3.4, 5.9].

\citet{Takahashi2010} developed a mass density dataset using measurements from the Space Environment Monitor instruments on the Geostationary Operational Environmental Satellites (GOES) satellites from 1980 through 1992, with most measurements in the range of $L=6.8\pm0.2 R_E$. The mass density was estimated using the Alfv\'en wave velocity relationship, $V_A=B/\sqrt{\mu_0\rho}$, a magnetic field model, and a numerical solution to a wave equation with an ionospheric boundary and the assumption of a zero resistance ionosphere.  The equatorial mass density, $\rho_{eq}$, was derived from the estimated mass density using a power law dependence on the geocentric distance to the field line of the observation, $R$, $\rho=\rho_{eq}(LR_E/R)^{1/2}$.  They found a high correlation ($\sim 0.94$) between 27-day averages of $F10.7$ and 27-day medians of $\rho_{eq}$.

\citet{Takahashi2010} noted that downward drops in the $D_{st}$ index coincided with significant changes in $\rho_{eq}$ for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ increases after $D_{st}$ minimum, two had $\rho_{eq}$ spikes before the minimum, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, the epoch average of $D_{st}$ for storms with a minimum $D_{st} < -50$~nT showed an enhancement in $\rho_{eq}$ the same day as minimum $D_{st}$ in the epoch averages.

\citet{Yao2008} studied the relationship between $D_{st}$ and the number density of low-energy $O^+$ in different regions (ring current and plasma sheet) using the TEAMS and ESA instruments on the low-altitude and high-inclination orbiting FAST satellite and found that the average $N_{O^+}$ across the sampled L-shells (2-14) had a strong correlation (0.88) with the minimum $D_{st}$ of the storm.  Although the correlation at geosynchronous distances was not calculated, the data presented for four events show that the enhancements in $N_{O^+}$ tended to appear near $D_{st}$ minimum ($\pm$ 2~hours).

The above results indicate that (1) the mass density at geosynchronous distances is best correlated with $F10.7$ and (2) there exists a much weaker statistical relationship between mass density and the geomagnetic activity indices $K_p$ and $D_{st}$.  In this work we consider the dependence of mass density estimates in the \citet{Takahashi2010} data set and attempt to identify and statistically characterize the relationship between geomagnetic activity with mass density at geosynchronous distances.  In addition, we attempt to identify solar wind variables that may drive or be associated with the changes.

%There are several issues that are addressed in detail: (1) the plasma density measurements are sparse, (2) The plasma trough density has a high correlation with F10.7, and geomagnetic and solar wind parameters also correlate with F10.7, (3) the dependence of mass density on the north/south component of the solar wind magnetic field is expected to be nonlinear. For northward IMF, corresponding to low solar wind driving of the magnetosphere, the plasmasphere can expand past geosynchronous orbit where trough estimates of mass density are made resulting in an increase in observed density that are not due to solar wind driving; for southward IMF, events studies have shown an increase in mass density associated with strong southward IMF and enhanced magnetospheric activity (\citet{Takahashi2010}; \citet{Yao2008}).

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ are from the dataset of \citet{Takahashi2010} data are available from 1980 through 1991 from GOES 2, 3, 5, 6 and 7.  All other parameters used are from \citet{Kondrashov2014ReconstructionOfGaps}, which has coverage from 1972 through 2013 and are on a 1--hour time grid. $\rho_{eq}$ is the inferred equatorial mass density based on the 3rd harmonic torodial frequency of magnetic field measurements as described in \citet{Takahashi2010}.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly median over the same interval for which the solar wind parameters were averaged, the median of all $\rho_{eq}$ values in a given hour window was used. GOES 6 had the longest span of available (May 1983 to August 1991) data and the primary results presented are for GOES 6 while measurements from other satellites were used for consistency checks.

@@

The top panel of Figure~\ref{fig:ccplot} shows the long-term trends of log$_{10}(\rho_{eq})$ from GOES 6 and $F_{10.7}$ computed using the median values in 27-day non-overlapping windows.  A scatter plot of these two lines is shown in the lower panel of Figure~\ref{fig:ccplot} along with that for 1--day and 1--hour medians.  The linear correlation for GOES 6 using all measurements is found to be $0.94$, which is the same value documented by \citet{Takahashi2010} who used combined measurements from all satellites in the time interval of 1980 through 1991 using the constraint $06\mbox{:}00 \leq MLT \le 12\mbox{:}00$, $K_{p3d}\ge1.0$ and $D_{st}\ge -50$~nT.  Our 27-day correlations for GOES 2, 5, and 7 are 0.81, 0.78, and 0.87, and the time range of coverage of measurements from these satellites is 1980-01-01/1983-05-16, 1983-01-01/1987-02-28, and 1983-05-27/1991-08-29, respectively.

@@

Two types of events are considered. The first is a drop in the $D_{st}$ index below the threshold value of -50~nT. These types of events are considered in order to determine if there is a well-defined $\rho_{eq}$ dependence on geomagnetic activity as indicated by the $D_{st}$ index.  It is expected that the plasmasphere will exhibit two responses near the time of these threshold crossings.  First is the movement of the plasmapause earthward (\citet{LemaireEarthsPlasmasphere}) and the second is a possible change in density due to magnetospheric processes.

@@

Our first analysis uses GOES-6 data only from the time interval 1989-1991, which corresponds to the interval used in the epoch analysis of \citet{Takahashi2010}. In this interval, there were 75 $D_{st}$ events where $D_{st}$ remained below the $-50$~nT threshold for at least 12 hours. Only events with an onset time when the spacecraft was located between 06:00 and 12:00 MLT were considered. The first hour of a local minimum in $D_{st}$ following a threshold crossing defines the zero epoch time, and for each event, $4\cdot24$ hours were considered before and $8\cdot24$ hours after onset. For multiple detected threshold crossings within 24 hours of each other, the first crossing was selected as the event onset because visual inspection showed the subsequent crossings to typically be from brief downward deviations during the recovery period of a geomagnetic storm. To compute the epoch averages on a daily time scale, the median value of all available measurements for all events centered on a window of $\pm 12$ hours of the epoch zero hour was computed, and these averaging windows were shifted in increments of $24$~hours. Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians for each event in 1-day bins and then compute the medians across the averaged event time series on each epoch day.

The stack plot in the upper panel of Figure \ref{fig:DailyAveragedDstEvents} shows the epoch averages for the 75 events from 1989-1991 using data from GOES 6.  The minimum $D_{st}$ median is $-75$~nT.  Consistent with \citet{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$ on the day of the event relative to the day before and after, although the magnitude of increase observed here is $4.5$ amu/cm$^3$ instead of the increase of ${\sim} 10$ amu/cm$^3$ found in Figure 11 of \citet{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute the medians and red error lines.  The uncertainty is shown with error lines for each parameter that have a width that is twice the standard deviation of the values used in computing the median divided by the square root of the number of values. For the non-$\rho_{eq}$\ parameters, the number of measurements used in computing the medians is typically equal to the number of events, which is much larger than the number of $\rho_{eq}$\ values as all available measurements were used instead of restricting to only values where a $\rho_{eq}$ value also existed.

@@

To increase the number of events, we have removed the constraint that the events must occur in the MLT range of 06:00-12:00 and the constraint that $D_{st}$ remained below the threshold value for 12 hours or longer.  The result is shown in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents}. The two variations from a baseline (one downwards before onset and one upwards after onset) were tested for significant deviation from two baselines, one 24-12 hours before onset, one 16-48 hours after onset. The probability of the downward variation coming from the same distribution as the data from the left baseline is 1.06\%, whereas the probability of the right upward variation coming from the same distribution of the data for the right baseline is 0.17\%.  This significant result is also maintained for GOES 2, 5, and 7.  This indicates that there are statistically significant variations in $\rho_{eq}$ within 6 hours of onset when hourly averages are considered.  The signficance of the upward variation result is consistent with previous observations of large mass density increases in the plasma trough region during large geomagnetic storms (\citet{Yao2008}; \citet{Takahashi2010}).

@@

Figure~\ref{fig:HourlyAveragedRhoEvents} can be compared with Figure~9 of \citet{Denton2016}, which shows daily averages of $\rho_{eq}$ from this dataset after the onset of quiet $K_p$ intervals.  The growth shown here (using the $\rho_m$ threshold criteria with no constraint on $K_p$) is approximately linear and lasts $\sim$ 12--18 hours whereas the 10 $K_p$ events (elevated $K_p$ followed by 2 days with $K_p \le 1^+$) considered by \citet{Denton2016} showed growth that lasted for at least 48~hours.

@@

Figure~\ref{fig:F107Binned} shows how the epoch average curve of $\rho_{eq}$ in Figure~\ref{fig:HourlyAveragedRhoEvents} depends on $F10.7$.  Statistically significant differences occur independent of whether the separating factor is the average of $F10.7$ before or after onset, as expected because the cadence of $F10.7$ is one day and its time scale of variation is longer than that of $\rho_{eq}$.  Events with high $F10.7$ are associated with higher $\rho_{eq}$ before and after event onset but similar peak values.  As was the case for $B_z$, there is an approximately 12--hour time span of linear growth of $\rho_{eq}$ near the time of onset but because the initial starting $\rho_{eq}$ is higher for high $F10.7$, the associated growth rate is lower, consistent with \citet{Denton2016}.

@@

In recent years, various works have considered the influence of solar wind control and the geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit, generally using event studies and linear correlations (\citet{Takahashi2006}; \citet{Denton2006}; \citet{Yao2008}; \citet{Takahashi2010}; \citet{Denton2011}; \citet{Denton2016}).  The event studies show a possible relationship between $\rho_{eq}$ and solar wind and geomagnetic activity indicators such as $K_p$ and $D_{st}$ while correlation studies have generally found a very weak relationship.

@@

%Based on a weak linear relationship between $\rho_{eq}$ and $B_z$, \citet{Denton2016} speculated that this may be a result of a closed magnetosphere being more conducive to plasmasphere refilling.

@@

For $\rho_{eq}$ events, more positive $B_{z}$ after onset corresponded to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may be due to the apparent mass refilling mechanism described by \citet{Denton2016}.  In order to separate this mechanism from an enhancement due to magnetospheric activity using this type of analysis, additional information about the location of the plasmapause may be required.

@@

\newpage
%\footnotesize
%\bibliographystyle{plainnat}
%\bibliography{paper}

\begin{thebibliography}{16}
	\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

	\bibitem[{{\it Carpenter\/}(1966)}]{Carpenter1966}
	Carpenter, D.~L., Whistler studies of the plasmapause in the magnetosphere: 1.
	{T}emporal variations in the position of the knee and some evidence on plasma
	motions near the knee, {\it J. Geophys. Res.\/}, {\it 71\/}, 693--709, 1966.

	\bibitem[{{\it Carpenter and Anderson\/}(1992)}]{Carpenter1992}
	Carpenter, D.~L., and R.~R. Anderson, An {ISEE}/whistler model of equatorial
	electron density in the magnetosphere, {\it J. Geophys. Res.\/}, {\it 97\/},
	1097--1108, 1992.

	\bibitem[{{\it Chappell et~al.\/}(1970)}]{Chappell1970}
	Chappell, C.~R., K.~K. Harris, and G.~W. Sharp, A study of the influence of
	magnetic activity on the location of the plasmapause as measured by {OGO} 5,
	{\it J. Geophys. Res.\/}, {\it 75\/}, 50--56, 1970.

	\bibitem[{{\it {Denton}\/}(2007)}]{Denton}
	{Denton}, R., {Database of Input Parameters for Tsyganenko Magnetic Field
		Models}, 2007.

	\bibitem[{{\it {Denton} et~al.\/}(2006)}]{Denton2006}
	{Denton}, R.~E., K.~{Takahashi}, I.~A. {Galkin}, P.~A. {Nsumei}, X.~{Huang},
	B.~W. {Reinisch}, R.~R. {Anderson}, M.~K. {Sleeper}, and W.~J. {Hughes},
	{Distribution of density along magnetospheric field lines}, {\it J. of
		Geophys. Res.\/}, {\it 111\/}, 4213, 2006.

	\bibitem[{{\it Denton et~al.\/}(2011)}]{Denton2011}
	Denton, R.~E., M.~F. Thomsen, K.~Takahashi, R.~R. Anderson, and H.~J. Singer,
	Solar cycle dependence of bulk ion composition at geosynchronous orbit, {\it
		J. Geophys. Res.\/}, {\it 116\/}, n/a--n/a, 2011.

	\bibitem[{{\it Denton et~al.\/}(2016)}]{Denton2016}
	Denton, R.~E., K.~Takahashi, J.~Amoh, and H.~J. Singer, Mass density at
	geostationary orbit and apparent mass refilling, {\it J. Geophys. Res.\/},
	{\it 121\/}, 2962--2975, 2016.

	\bibitem[{{\it {Gallagher} et~al.\/}(1988)}]{Gallagher1988EmpiricalModelPlasmasphere}
	{Gallagher}, D.~L., P.~D. {Craven}, and R.~H. {Comfort}, {An empirical model of
		the earth's plasmasphere}, {\it Advances in Space Research\/}, {\it 8\/},
	15--21, 1988.

	\bibitem[{{\it {Kondrashov} et~al.\/}(2014)}]{Kondrashov2014ReconstructionOfGaps}
	{Kondrashov}, D., R.~{Denton}, Y.~Y. {Shprits}, and H.~J. {Singer},
	{Reconstruction of gaps in the past history of solar wind parameters}, {\it
		Geophys. Res. Lett.\/}, {\it 41\/}, 2702--2707, 2014.

	\bibitem[{{\it Lemaire\/}(1998)}]{LemaireEarthsPlasmasphere}
	Lemaire, J., {\it The Earth's Plasmasphere\/}, Cambridge University Press,
	1998.

	\bibitem[{{\it Maynard and Grebowsky\/}(1977)}]{Maynard1977}
	Maynard, N.~C., and J.~M. Grebowsky, The plasmapause revisited, {\it J.
		Geophys. Res.\/}, {\it 82\/}, 1591--1600, 1977.

	\bibitem[{{\it {Moldwin} et~al.\/}(2002)}]{Moldwin2002ModelPlasmapause}
	{Moldwin}, M.~B., L.~{Downward}, H.~K. {Rassoul}, R.~{Amin}, and R.~R.
	{Anderson}, {A new model of the location of the plasmapause: CRRES results},
	{\it J. of Geophys. Res.\/}, {\it 107\/}, 1339, 2002.

	\bibitem[{{\it {O'Brien} and
			{Moldwin}\/}(2003)}]{OBrien2003EmpiricalPlasmapause}
	{O'Brien}, T.~P., and M.~B. {Moldwin}, {Empirical plasmapause models from
		magnetic indices}, {\it Geophys. Res. Lett.\/}, {\it 30\/}, 1--1, 2003.

	\bibitem[{{\it {Takahashi} et~al.\/}(2006)}]{Takahashi2006}
	{Takahashi}, K., R.~E. {Denton}, R.~R. {Anderson}, and W.~J. {Hughes}, {Mass
		density inferred from toroidal wave frequencies and its comparison to
		electron density}, {\it J. of Geophys. Res.\/}, {\it 111\/}, 1201, 2006.

	\bibitem[{{\it {Takahashi} et~al.\/}(2010)}]{Takahashi2010}
	{Takahashi}, K., R.~E. {Denton}, and H.~J. {Singer}, {Solar cycle variation of
		geosynchronous plasma mass density derived from the frequency of standing
		Alfv{\'e}n waves}, {\it J. of Geophys. Res.\/}, {\it 115\/}, 7207, 2010.

	\bibitem[{{\it {Yao} et~al.\/}(2008)}]{Yao2008}
	{Yao}, Y., K.~{Seki}, Y.~{Miyoshi}, J.~P. {McFadden}, E.~J. {Lund}, and C.~W.
	{Carlson}, {Effect of solar wind variation on low-energy O$^+$ populations in
		the magnetosphere during geomagnetic storms: FAST observations}, {\it J. of
		Geophys. Res.\/}, {\it 113\/}, 2008.

\end{thebibliography}

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.45]{UsedFigures/2016SW001507R-p01.eps}
\caption{Overview of data used in this article. The top four panels show parameters from \citet{Kondrashov2014ReconstructionOfGaps} and the bottom panel contains $\rho_{eq}$ based on GOES 6 measurements from \citet{Denton} after interpolation and averaging described in the text. Dashed horizontal lines in the $D_{st}$ and $\rho_{eq}$ panels indicate sample event cutoff thresholds of $D_{st} = -50$~nT and $\rho_{eq} = 20$~amu/cm$^3$ considered in Section~3.}
\label{fig:AllData}
\end{figure}
